Sequence Labeling for Gait Analysis using LSTM

### 2015-05-09 paper/paper.tex

\section{Data}
\label{sec:Data}

Our training and testing was conducted on data from 10 healthy subjects walking at three different speeds for 50 seconds. Each sequence of 50 seconds will be called a trial from now on. This totals 1500 seconds of recorded data, which roughly correspond to 1000 gait cycles.  For each subject the motion-capture data (input) is 3D marker position of 18 different markers placed in the same anatomical positions for each subject, sampled at 100Hz. The event labels (target) were computed from ground force reaction data sampled at 1kHz, which is considered the gold standard in the field. In order to compute events the ground reaction forces in gravity's axis are thresholded ($20 N$) and whenever they exceed this threshold, the foot is considered to be on the ground (stance). By searching for the transitions between stance and no-stance, the events are found.

The markers were placed at the toe, ankle, heel, knee, greater trochanter (hip), tibia, femur, and anterior and posterior superior illiac spine (pelvis) on each leg.
% All data was collected in the Human Movement Research Laboratory of the University of Pittsburgh, with approval of its IRB. All subjects consented to allow their data be used for research purposes. We are using the data with permission from the laboratory's PI, Dr. Gelsy Torres-Oviedo.

The available data was separated into a training and a testing set in two different ways. Dataset 1 takes the first 25 seconds of data from each trial as the training sample, and the following 25 seconds as the testing sample. This dataset was meant to test how well a trained network can generalize to new gait cycles, given that it was trained on data including the same subject and walking speed. As such, it is the most basic test of performance that can be done. Dataset 2 includes the same data, but splits it along trials: all 50 seconds of 15 trials were used for training, and the remaining 15 trials were used for testing. It is worth noticing that this second dataset necessarily has different subjects in its training and testing sets, so it allows us to assess how the algorithm generalizes to new users without being trained on them.

>>>>

\section{Data}
\label{sec:Data}
Our training and testing was conducted on data from 10 healthy subjects walking 
on a treadmill at three different speeds for 50 seconds. Each sequence of 50 seconds is called a trial. This amounts to a total of 1500 seconds of recorded data, which roughly correspond to 1000 gait cycles.  For each subject, 3D motion capture data was collected using 18 reflective markers placed at precise body locations.%
\footnote
{
    The locations used were toe, ankle, heel, knee, greater trochanter (hip), tibia, femur, and anterior and posterior superior illiac spine (pelvis) on each leg.
}
The sampling frequency of this data is 100Hz. The event labels (target) were computed from ground force reaction data from force sensors embedded in the treadmill 
which is considered the gold standard in the field. 
In order to compute events the ground reaction forces in gravity's axis are thresholded ($20 N$) and whenever they exceed this threshold, the foot is considered to be on the ground (stance). By searching for the transitions between stance and no-stance, the events are found.
The sampling frequency of this data is 1kHz.

The data was separated into a training and a testing set in two different ways. Dataset 1 takes the first 25 seconds of data from each trial as the training sample, and the following 25 seconds as the testing sample. This dataset was meant to test how well a trained network can generalize to new gait cycles, given that it was trained on data including the same subject and walking speed. As such, it is the most basic test of performance that can be done. Dataset 2 includes the same data, but splits it along trials: all 50 seconds of 15 trials were used for training, and the remaining 15 trials were used for testing. It is worth noticing that this second dataset necessarily has different subjects in its training and testing sets, so it allows us to assess how the algorithm generalizes to new users without being trained on them.
